Accelerating Deep Neural Network Training on FPGAs Facilitating the Use of Variable Precision

### 2018-06-29 SWoPP2018 Paper/paper.tex
Took out the out of order loops and wrote some more for the paper.

\documentclass[techrep,english]{ipsj}
\usepackage[utf8]{inputenc}

\usepackage[dvips]{graphicx}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{adjustbox}
\usepackage{booktabs}
\usepackage{siunitx}

@@

@@

%\received{2011}{7}{1}
%\rereceived{2011}{10}{1}   % optiona
%\rerereceived{2011}{10}{31} % optional
%\accepted{2011}{11}{5}

@@

@@

%1
\section{Introduction}
The present paper describes our efforts toward a customized and generalized framework for design-space exploration using alternative numerical floating-point formats through FPGAs.

The recent explosion in artificial intelligence – in particular that of deep neural networks based on backpropagation – has triggered a storm in the introduction and creation of specialized compute devices. Commercial platforms such as Microsoft’s BrainWave~\cite{msbrainwave}, Google’s TPUs~\cite{googletpu}, and Fujitsu’s DLU~\cite{fujitsudlu} are all examples of specialized circuitry dedicated to low-power, high-performance Deep-Learning (DL) training and/or inference. Meanwhile, existing general-purpose manufacturers are empowering their architecture with custom floating-point units that trade precision for performance, such as Intel’s Knights-Mill\`\cite{knm} and NVIDIA’s Volta-100\`\cite{volta100}. It is clear that Artifical Intelligence and Deep-Learning will have a prioritized presence in modern architecture -- today and in the near future.

@@

The remainder of our paper is structure in the following way: Section~\ref{sec:relwork} position our work against other, similar efforts. Section~\ref{sec:framework} gives a background to machine learning and FPGAs, and describes our proposed architecture. Section~\ref{sec:method} overviews our experimental methodology, and is followed by our preliminary results in Section\`\ref{sec:result}. We conclude in section~\ref{sec:conclusion}.

@@

\section{A FPGA-based CNN Framework}\label{sec:framework}
\subsection{Field-Programmable Gate-Arrays}
%Small intro to FPGAs.
FPGAs can be used for a great variety of things, from accelerating computational algorithms to implementating a hardware design before taping out full chip.
They are devices that can represent reconfigurable hardware. 
The fabric they are made up of consists of vast arrays of switches and other routing hardware to connect all the functional elements, like multipliers, block RAM and look up tables, in any way possible.
This makes them very suitable for the research described in this paper.
It gives total freedom to design a hardware implementation to explore algorithms in a hardware setting.

\subsection{Neural Networks}
Intro to neural networks.

\subsection{FPGA NN Framework}
Small intro to the framework.
\subsubsection{Design overview}
Overview design + picture.
\subsubsection{Design details}
All weight matrices are transposed in memory.
The main inference is done using an blocking implementation of the matrix multiplication.
So a block is fetched from the activations and the wieghts and place in the block RAM on the FPGA.
The block RAM is much faster then the global memory.
This speeds up through put by a large margin.
Block RAM is equivalent to the shared or local memory of contemporary GPUs.

@@

\section{Methodology}\label{sec:method}
\subsection{Experimental Platform}
\begin{table}[h]\label{tab:software-versions}
  \caption{Used software versions}
  \begin{tabular}{ll}
    \toprule
    \textbf{Software} & \textbf{Version} \\
    \midrule
    Python & v3.6.3 \\
    Keras & v2.2.0 \\
    TensorFlow & v1.8.0 \\
    Intel Quartus & v17.1.0 \\
    \bottomrule
  \end{tabular}
\end{table}
\begin{table}[h]\label{tab:test-hardware}
  \caption{Test hardware}
  \begin{tabular}{ll}
    \toprule
    \textbf{Hardware} & \textbf{Configuration} \\
    \midrule
    Intel i7-3930K & 3.20 GHz \\
    Arria 10 & Nallatech N510T \\
    TensorFlow & v1.8.0 \\
    Intel Quartus & v17.1.0 \\
    \bottomrule
  \end{tabular}
\end{table}
Compiler, version, flags, fpga boards, host, etc.
\subsection{Network Tested + Data-set}
The used example network is a simple multilayer perceptron.
The main dataset used was MNIST with a layer depth of 1 and the first layer output size of 128.
The data set was trained using the first \num{50000} images from the MNIST dataset.
The last \num{10000} images are used as the test dataset.
Training was done using a batch size of 128.
The CPU code was implemented using the NumPy array.dot function.

\section{Results}\label{sec:result}

\subsection{FPGA Resource Utilization}
Area, DSP, Fmax, etc.

\subsection{Training Performance}
Convergence, performance, accuracy, etc.

>>>>

\documentclass[techrep,english]{ipsj} % techrep,

  \usepackage[nocompress]{cite}

  %\usepackage[dvips]{graphicx}
  \usepackage{latexsym}
  \usepackage{url}

  \usepackage{diagbox}

  \usepackage{tabu}
  \usepackage{footnote}
  \usepackage{tablefootnote}
  \usepackage{enumitem}
  \usepackage{times}
  \usepackage{float}
  \usepackage[binary-units=true,quotient-mode=fraction,table-auto-round]{siunitx}
  \sisetup{detect-all}
  \sisetup{range-phrase=--}
  \sisetup{range-units=single}
  \DeclareSIUnit\years{years}
  \usepackage{graphicx}
  \usepackage{subfigure}
  \usepackage{amsmath,amsfonts,amssymb}
  \usepackage{comment}
  \usepackage{multirow}
  \usepackage{bbding}
  \newcommand{\tabincell}[2]{\begin{tabular}{@{}#1@{}}#2\end{tabular}}
  \usepackage{times}
  \usepackage{soul}
  \usepackage{color}
  \usepackage{latexsym}
  \usepackage{booktabs,multirow,multicol}
  \usepackage{todonotes}
  \usepackage{multirow}
  \usepackage{comment}
  \usepackage{tabularx}
  \usepackage{url}

@@

\graphicspath{
    {resources/}
}

%\def\Underline{\setbox0\hbox\bgroup\let\\\endUnderline}
%\def\endUnderline{\vphantom{y}\egroup\smash{\underline{\box0}}\\}
%\def\|{\verb|}

@@

@@

\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink,noabbrev,capitalise]{cleveref}

@@

\section{Introduction}
The present paper describes our efforts toward a customized and generalized framework for design-space exploration using alternative numerical floating-point formats through FPGAs.

The recent explosion in artificial intelligence – in particular that of deep neural networks based on backpropagation – has triggered a storm in the introduction and creation of specialized compute devices.
Commercial platforms such as Microsoft’s BrainWave~\cite{msbrainwave}, Google’s TPUs~\cite{googletpu}, and Fujitsu’s DLU~\cite{fujitsudlu} are all examples of specialized circuitry dedicated to low-power, high-performance Deep-Learning (DL) training and/or inference.
Meanwhile, existing general-purpose manufacturers are empowering their architecture with custom floating-point units that trade precision for performance, such as Intel’s Knights-Mill~\cite{knm} and NVIDIA’s Volta-100~\cite{volta100}.
It is clear that Artifical Intelligence and Deep-Learning will have a prioritized presence in modern architecture -- today and in the near future.

@@

The remainder of our paper is structure in the following way: \cref{sec:relwork} position our work against other, similar efforts.
\cref{sec:framework} gives a background to machine learning and FPGAs, and describes our proposed architecture.
\cref{sec:method} overviews our experimental methodology, and is followed by our preliminary results in \cref{sec:result}.
We conclude in \cref{sec:conclusion}.

@@

\section{A FPGA-based CNN Framework}\label{sec:framework}
\subsection{Field-Programmable Gate-Arrays}
%Small intro to FPGAs.
FPGAs can be used for a great variety of things, from accelerating computational algorithms to implementating a hardware design before taping out full chip.
They are devices that can represent reconfigurable hardware.

The fabric they are made up of consists of vast arrays of switches and other routing hardware to connect all the functional elements, like multipliers, block RAM and look up tables, in any way possible.
This makes them very suitable for the research described in this paper.
It gives total freedom to design a hardware implementation to explore algorithms in a hardware setting.

\subsection{Neural Networks}
Intro to neural networks.
%TODO

\subsection{FPGA NN Framework}
A small framework was built, consisting of multiple kernels.
One simple forward propagation kernel with optional Relu activation function integrated and a forward softmax kernel.
\subsubsection{Design overview}
\Cref{fig:block-diagram} shows a overview of the system.
The backwards kernels are a work in progress, so for now we use an external training script.
The host code, written in python, does a kernel call for every layer.

\begin{figure*}[ht]
  \centering
  \includegraphics[width=\linewidth]{block-diagram.pdf}
  \caption{Block diagram of implementation}\label{fig:block-diagram}
\end{figure*}

\subsubsection{Design details}
All weight matrices are transposed in memory.
The main inference is done using an blocking implementation of the matrix multiplication.
So a block is fetched from the activations and the wieghts and place in the block RAM on the FPGA.
The block RAM is much faster then the global memory.
This speeds up through put by a large margin.
Block RAM is equivalent to the shared or local memory of contemporary GPUs.

@@

\section{Methodology}\label{sec:method}
\subsection{Experimental Platform}
The software versions and hardware configuration are shown in respectively \cref{tab:software-versions} and \cref{tab:test-hardware}.
All kernels were compiled with \texttt{-O3 -fp-relaxed}.
\texttt{-fp-relaxed} gives the compiler some more freedom with the order of operations, this gives about a \SIrange{10}{15}{\percent} performance benefit without changing the numerical output.
\begin{table}[h]
  \centering
  \caption{Used software versions}\label{tab:software-versions}
  \begin{tabular}{ll}
    \toprule
    \textbf{Software} & \textbf{Version} \\
    \midrule
    Python & v3.6.3 \\
    Keras & v2.2.0 \\
    TensorFlow & v1.8.0 \\
    Intel Quartus & v17.1.0 \\
    \bottomrule
  \end{tabular}
\end{table}
\begin{table}[h]
  \centering
  \caption{Test hardware}\label{tab:test-hardware}
  \begin{tabular}{ll}
    \toprule
    \textbf{Hardware} & \textbf{Configuration} \\
    \midrule
    Intel i7-3930K & \SI{3.20}{\giga\hertz} \\
    Host RAM & DDR3-\SI{1333}{\mega\hertz} Dual Channel \\
    Arria 10 (x1) & Nallatech 510T \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Network Tested + Data-set}
The used example network is a simple multilayer perceptron.
The main dataset used was MNIST with a layer depth of 1 and a first layer output size of 128.
So in total a two layer network.
The data set was trained using the first \num{50000} images from the MNIST dataset.
The last \num{10000} images are used as the test data set.
Training was done using a batch size of 128.
The CPU code was implemented using the NumPy array.dot function.

\section{Results}\label{sec:result}
The host (numpy based) code benchmarked at \num{41} GFLOPS for simple inference.
Our first naive implementation clocked about \num{2.5} GFLOPS for this problem size.
The results for the different final FPGA kernels are shown in \cref{tab:kernel-performance}.

\begin{table}[h]
  \centering
  \caption{FPGA Kernel performance}\label{tab:kernel-performance}
  \begin{tabular}{ll}
    \toprule
    \textbf{Block size} & \textbf{GFLOPS}  \\
    \midrule
    128x128 & \num{0}  \\
    100x100 & \num{0} \\
    84x84 & \num{0}  \\
    64x64 & \num{14.3}  \\
    32x32 & \num{0}  \\
    16x16 & \num{4.1}  \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{FPGA Resource Utilization}
The kernel resource utilization is shown in \cref{tab:fpga-util}.

\begin{table}[h]
  \centering
  \caption{FPGA Hardware utilisation for different blocksizes}\label{tab:fpga-util}
  \begin{tabular}{lllll}
    \toprule
    \textbf{Block size} & \textbf{Logic} & \textbf{DSPs} & \textbf{RAMs} & \textbf{Fmax} \\
    \midrule
    128x128 & \SI{0}{\percent} & \SI{0}{\percent} & \SI{0}{\percent} & \SI{0}{\mega\hertz}  \\
    100x100 & \SI{0}{\percent} & \SI{0}{\percent} & \SI{0}{\percent} & \SI{0}{\mega\hertz}  \\
    84x84 & \SI{0}{\percent} & \SI{0}{\percent} & \SI{0}{\percent} & \SI{0}{\mega\hertz}  \\
    64x64 & \SI{33}{\percent} & \SI{10}{\percent} & \SI{68}{\percent} & \SI{172.8}{\mega\hertz}  \\
    32x32 & \SI{0}{\percent} & \SI{0}{\percent} & \SI{0}{\percent} & \SI{0}{\mega\hertz}  \\
    16x16 & \SI{17}{\percent} & \SI{3}{\percent} & \SI{30}{\percent} & \SI{209.16}{\mega\hertz}  \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Training Performance}
The accuracy after training on the CPU using Keras is \SI{97.9}{\percent}.
The FPGA gets slightly different floaing point results due to rounding differences but is matches the accuracy.
